---
title: "NES-LTER Utility Functions: Finding Paired Casts"
author: "Jaxine Wolfe"
date: "13 December 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compile Cruise Data from REST API

**Description:** read_from_api compiles CTD, nutrient, or chlorophyll data from the [WHOI REST API](https://github.com/WHOIGit/nes-lter-ims/wiki/Using-REST-API) for the given cruises. NOTE: this function is required to use the find_paired_casts function.

**Arguments:** 
    type         specify the type of data to fetch (options: "metadata", "summary", "nutrient", "chl", "bottles")
    cruises      a vector of cruises for which the data will be compiled (single or multiple cruises accepted)

```{r}

read_from_api <- function(type, cruises) {
  # expand the cruises into a dataframe (avoids nested for loops)
  z <- expand.grid(cruises)
  
  # read in data based on the specified source
  if (type == "metadata") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/metadata.csv")
    urls <- unlist(urls)
  }
  if (type == "summary") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/bottle_summary.csv")
    urls <- unlist(urls)
  }
  if (type == "nutrient") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/nut/", z$Var1, ".csv")
    urls <- unlist(urls)
  }
  if (type == "chl") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/chl/", z$Var1, ".csv")
    urls <- unlist(urls)
  }
  if (type == "bottles") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/bottles.csv")
    urls <- unlist(urls)
  }

  ## Cruise Compilation ##
  # case: more than one cruise given
  if (length(cruises) > 1) {
  # begin compilation  
  prev_cruise <- read_csv(urls[1])
  
  # loop through urls to compile cruise data into one file
  for (k in 2:length(urls)){
    # read in data per cruise
    next_cruise <- read_csv(urls[k])
    
    # bind the next cruise to the compiled cruise dataset
    all <- plyr::rbind.fill(prev_cruise, next_cruise)
    
    # if statment to reset the previous cruises until all cruises are read in
    if(k < length(urls)) {
      prev_cruise <- all
    }
  }
  return(all)

  # case: only one cruise is given
  } else {
   all <- read_csv(urls)
   return(all)
  }
}

```

## Find Paired Casts

**Description:** find_paired_casts returns the provided data frame with an added field that indicates an associated paired cast if one exists. 

**Required:** 
  • tidyverse package
  • read_from_api function to access CTD data

**Arguments:** 
    df           a data frame (must contain columns cruise, cast, and date)
    min.time     the window of minimum time difference (in mins) within paired casts should be found

```{r}

find_paired_casts <- function(df, min.time) {
  # extract cruises from data
  cruiselist <- unique(df$cruise)
  # read in ctd data
  ctd_metadata <- read_from_api(type = "metadata", cruises = cruiselist)
  ctd_summary <- read_from_api(type = "summary", cruises = cruiselist)
  
  # define column to display paired cast
  df$paired_cast <- NA_character_
  
  for (j in 1:nrow(df)) {
    # store values for current row
    df_cruise <- df$cruise[j]
    df_cast <- df$cast[j]
    
    # isolate cast metadata and bottle summary
    cast_metadata <- ctd_metadata %>% filter(cruise == df_cruise & cast == df_cast)
    cast_summary <- ctd_summary %>% filter(cruise == df_cruise & cast == df_cast)
    
    # define the min and max time for the cast 
    df_date_max <- max(cast_summary$date)
    df_date_min <- min(cast_summary$date)
    
    # find the cast before from the bottle summary
    cast_before <- ctd_summary %>% 
      filter(cruise == df_cruise & cast == (df_cast-1))
    # find the cast after from the ctd metadata
    cast_after <- ctd_metadata %>% 
      filter(cruise == df_cruise & cast == (df_cast+1))
    
    # deterine the time difference
    time_before <- min(abs(as.numeric(difftime(df_date_min, cast_before$date, units = "mins"))),
                       na.rm = TRUE)
    time_after <- min(abs(as.numeric(difftime(df_date_max, cast_after$date, units = "mins"))),
                      na.rm = TRUE)
    
    # determine paired casts if they exist
    if (time_after > min.time & time_before > min.time) {
      j <- j + 1
      next
    }
    if (time_before < min.time) {
      # isolate cast metadata to extract station
      before_metadata <- ctd_metadata %>% filter(cruise == df_cruise & cast == (df_cast-1))
      if (isFALSE(cast_metadata$nearest_station == before_metadata$nearest_station)) {
        print(paste0("paired cast found for cruise ", df_cruise, 
                     " and cast ", df_cast, ", but nearest station did not match"))
        j <- j + 1
        next
      } else {
        df$paired_cast[j] <- df_cast - 1
      }
    }
    if (time_after < min.time) {
      if (isFALSE(cast_metadata$nearest_station == cast_after$nearest_station)) {
        print(paste0("paired cast found for cruise ", df_cruise, 
                     " and cast ", df_cast, ", but nearest station did not match"))
        j <- j + 1
        next
      } else {
        df$paired_cast[j] <- df_cast + 1
      }
    }
  }
  return(df)
}

```

