---
title: "NES-LTER Utility Functions: Compiling Cruise Data from WHOI REST API"
author: "Jaxine Wolfe"
date: "13 December 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compile Cruise Data from REST API

**Description:** read_from_api compiles CTD, nutrient, or chlorophyll data from the [WHOI REST API](https://github.com/WHOIGit/nes-lter-ims/wiki/Using-REST-API) for the given cruises. 

*Types of data this function can acquire:*
  • metadata: CTD metadata provides a list of casts for a given cruise along with metadata about each cast (e.g., time, lat/lon, nearest station)
  • summary: CTD bottle summary data provides a concise summary of when and where each bottle was fired.
  • bottles: CTD bottle data provides a list of niskins along with data from the CTD as it has been provided by the SeaBird CTD processing software.
  • nutrient: Nutrient data provided per cruise.
  • chl: Chlorophyll data provided per cruise.

**Arguments:** 
    type         the type of data to fetch (options: "metadata", "summary", "nutrient", "chl", "bottles")
    cruises      a vector of cruises for which the data will be compiled (single or multiple cruises accepted)

```{r}

read_from_api <- function(type, cruises) {
  # expand the cruises into a dataframe (avoids nested for loops)
  z <- expand.grid(cruises)
  
  # read in data based on the specified source
  if (type == "metadata") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/metadata.csv")
    urls <- unlist(urls)
  }
  if (type == "summary") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/bottle_summary.csv")
    urls <- unlist(urls)
  }
  if (type == "nutrient") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/nut/", z$Var1, ".csv")
    urls <- unlist(urls)
  }
  if (type == "chl") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/chl/", z$Var1, ".csv")
    urls <- unlist(urls)
  }
  if (type == "bottles") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/bottles.csv")
    urls <- unlist(urls)
  }

  ## Cruise Compilation ##
  # case: more than one cruise given
  if (length(cruises) > 1) {
  # begin compilation  
  prev_cruise <- read_csv(urls[1])
  
  # loop through urls to compile cruise data into one file
  for (k in 2:length(urls)){
    # read in data per cruise
    next_cruise <- read_csv(urls[k])
    
    # bind the next cruise to the compiled cruise dataset
    all <- plyr::rbind.fill(prev_cruise, next_cruise)
    
    # if statment to reset the previous cruises until all cruises are read in
    if(k < length(urls)) {
      prev_cruise <- all
    }
  }
  return(all)

  # case: only one cruise is given
  } else {
   all <- read_csv(urls)
   return(all)
  }
}

```